---
title: 'coursera notes'
author: 'Jet'
output: pdf_document
---

# notes from r programming week 1 #
R has five basic or atomic classes of objects:
* character
* numeric
* integer
* logical (True/False)

The most basic object is a vector
* a vector can **only** contain objects of the same class
* list is the exception that is represented as a vector but can contain certain objects of different classes

True and False can be shortened to T and F 

A common way to create a matrix in r is by column binding or row binding vectors together into a matrix.

```{r}
x <- 1:3
y <- 10:12
cbind(x,y)
rbind(x,y)
```
# Factors #
factors contain ordered variables, and underlying them is the a labelled integer vector

```{r}
x <- factor(c("yes","yes", "no", "yes", "no"))
x
table(x)
unclass(x)
```
# Missing values #
NA is a missing value
NaN is not a number
NaN is a missing value but the reverse is not true

# reading in larger datasets with read.table
* check if your computer has enough memory to load the dataset
* if you know what type of data is in each column you can speed up the process
* a simple way to do this is to read the first 100 rows of the table, get the classes of each column using sapply, and then set the column classes based on that output, like so:
```{r}
initial <- read.table("datatable.txt", nrows = 100)
classes <- sapply (initial, class)
tabAll <- read.table("datatable.txt", colClasses = classes)
```




